Computing in the Field of Rationals

### 2012-06-24 rational_ext.tex
_

@@

\end{enumerate}
\end{proposition}
\begin{proof}
  We briefly summarize the execution of the first combinator (let us
  call it {{c}}) combinator to show that the coherence conditions
  hold. The second combinator simply mirrors {{c}}. To show that
  {{c=id}}, we must show for {{v:b}}, the evaluation of {{<c, v, []>}}
  always produces {{v:b}} as output.

@@

For a monoidal category to be compact closed the maps~{{eta}}
and~{{eps}} must satisfy a coherence condition that is usually
visualized as follows:
\begin{center}
  \includegraphics{diagrams/coherence.pdf}
\end{center}
where $b^*$ represents the dual of {{b}}.  In the case of negatives,
the condition amounts to checking that reversing direction twice is a
no-op. In the case of fractionals, the condition amounts to checking
that creating values {{alpha}} and {{1/alpha}} and immediately
unifying them is also a no-op. Both checks are straightforward and are
essentially the constructions in the introduction.

>>>>

\newcommand{\dgm}[2][0.95]{
\begin{center}
\scalebox{#1}{
\includegraphics{diagrams/#2.pdf}
}
\end{center}
}

@@

\end{enumerate}
\end{proposition}
\begin{proof}
  For convenience let us draw {{swap* (;) eps*}} shown on the left
  instead as shown on the right.

\begin{multicols}{2}
\dgm[0.8]{swap_eps_times}

\dgm[0.8]{cat_eps_times}    
\end{multicols}

The two coherence conditions are more intuitive when presented
graphically. They require that the combinators below be equal to
{{id}}, at the appropriate type (see Sec. 4
\cite{springerlink:10.1007/978-3-642-12821-94}).

\begin{multicols}{2}
\dgm{z_times1}

\dgm{z_times2}
\end{multicols}

  We briefly trace the execution of the first combinator (let us call
  it {{c}}. To show that {{c=id}}, we must show for {{v:b}}, the
  evaluation of {{<c,v,[]>}} always produces {{v:b}} as
  output. Verification for the second combinator is similar.

@@

It is important to note that in a compact closed category, the
dualizing action, usually denoted $(-)^\asterix$, is a contravariant
functor. However {{PiEE*}} is not compact closed and {{1/(-)}} is not
a functor specifically because divsion is not defined on 0 sized
types.

Without the important side condition on division by zero, we would
have {{1<->1/0*0}} which implies {{1 <-> 0}} and consequently
collapses the entire structure. Structures that are ``zero totalized
fields'' such as meadows \cite{Bergstra:2009:MES:1507774.1508000}
deserve investigation as alternative models for {{Pi}} with
fractionals (and negatives).

%% For a monoidal category to be compact closed the maps~{{eta}}
%% and~{{eps}} must satisfy a coherence condition that is usually
%% visualized as follows:
%% \begin{center}
%%   \includegraphics{diagrams/coherence.pdf}
%% \end{center}
%% where $b^*$ represents the dual of {{b}}.  In the case of negatives,
%% the condition amounts to checking that reversing direction twice is a
%% no-op. In the case of fractionals, the condition amounts to checking
%% that creating values {{alpha}} and {{1/alpha}} and immediately
%% unifying them is also a no-op. Both checks are straightforward and are
%% essentially the constructions in the introduction.
